NGO Repression as a Predictor of Worsening Human Rights Abuses

### 2021-11-19 analysis/index.Rmd
Add thing about PTS ceiling effect
Remove all civlib and clphy models

---
title: "Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
---

@@

Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. While research on democratic backsliding has tended to focus on rights and institutions in the electoral arena, this new wave of laws—many of which appear mundane on their face—represent a bureaucratic form of repression that could indicate more severe human rights abuses and backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. Do legal crackdowns on NGOs predict broader civil society repression? We argue they do, given that anti-NGO laws are among the most subtle means of repression and attract the least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of attacks on journalists, independent media, academic freedom, and freedoms of expression and association. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society.

>>>>

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
---

@@

An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while de jure anti-NGO laws provide little information in predicting future repression, their patterns of implementation—or de facto civil society repression—predicts worsening respect for physical integrity rights and civil liberties.

### 2021-11-19 manuscript/appendix.Rmd
Remove all civlib and clphy models

---
title: 'Appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 16, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
reference-section-title: References
bibliography: bibliography.bib
toc: false
appendix: true
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

>>>>

---
title: 'Appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
reference-section-title: References
bibliography: bibliography.bib
toc: false
appendix: true
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

if (is_docx) {
  library(huxtable)
}

### 2021-11-19 manuscript/manuscript.Rmd
Remove all civlib and clphy models

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 15, 2021"
published: "Conditionally accepted at the *Journal of Human Rights*"
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while de jure anti-NGO laws provide little information in predicting future repression, their patterns of implementation—or de facto civil society repression—predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "8,007"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

@@

  canary <- tar_read(panel)
  tar_load(vdem_clean)
  tar_load(civicus_map_data)
  tar_load(civicus_clean)

  tar_load(c(m_pts_total, m_pts_advocacy, m_pts_entry, m_pts_funding,
             m_lhr_total, m_lhr_advocacy, m_lhr_entry, m_lhr_funding,
             m_pts_v2csreprss, m_lhr_v2csreprss))

  # Model tables
  tar_load(c(models_paper_pts, models_paper_lhr))

  # Load lookup list for coefficients in regression tables
  tar_load(coef_list)

  # Load marginal effects
  tar_load(c(mfx_e1a_pts, mfx_e1d_lhr, mfx_e2a_pts, mfx_e2d_lhr))
})
```

@@

```{r civicus-map, fig.cap="2020 CIVICUS Monitor civic space ratings", fig.pos="h"}
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_viridis_d(option = "rocket", direction = -1,
                       na.translate = FALSE, name = "Civic space") +
  theme_void(base_size = 9, base_family = "Inter Light") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

@@

Simply noting the passage of (or counting) these laws may therefore miss the effect of their de facto implementation. Some governments may even pass laws to create a chilling effect on these organizations. By providing a legal threat, they could compel these organizations to self-censor without even needing to implement these laws in any systematic or meaningful capacity. These laws, especially during moments or crises or contention, "highlight to the wider NGO community-and citizenry-that the government is willing and capable of using the law to repress dissent" [@Brechenmacher:2017, 94].

@@

```{r repression, fig.width=5.5, fig.height=4, fig.asp=NULL, fig.cap="Key measures of repression over time"}
pts_plot <- canary %>% 
  filter(!is.na(PTS_factor)) %>% 
  group_by(PTS_factor, year) %>% 
  summarize(count = n()) %>% 
  mutate(fake_facet_title = "Political Terror Scale") %>% 
  ggplot(aes(x = year, y = count, color = PTS_factor)) +
  geom_line(size = 1) +
  scale_color_viridis_d(option = "rocket", end = 0.85, direction = -1) +
  coord_cartesian(xlim = c(1990, 2015)) +
  guides(color = guide_legend(title = NULL)) +
  labs(y = "Number of countries", x = NULL, caption = "(Higher PTS levels = more repression)") +
  theme_ngo() +
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines"),
        plot.caption = element_text(margin = margin(t = -1, unit = "lines"))) +
  facet_wrap(vars(fake_facet_title))

@@

```{r plot-mfx-e1a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on the probability of specific levels of political terror"}
mfx_e1a_pts %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  # Collapse 3-5 to just one level, since they're all tiny
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(x_rounded, effect2__, plot_var_nice) %>% 
  slice(1) %>% 
  # Add up these values for the collapsed 3-5 results
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = effect2__)) +
  geom_line(aes(group = effect2__), size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  # position = position_dodge(width = 0.2),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.85, direction = -1) +
  guides(color = guide_legend(override.aes = list(size = 0.5, linetype = 0))) +
  labs(x = "Count of legal barriers\n",
       y = "Predicted probability of category",
       color = "Predicted Political Terror Scale", fill = "Predicted Political Terror Scale",
       title = "Political Terror Scale") + 
  facet_wrap(vars(plot_var_nice), scales = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())
```

@@

```{r plot-mfx-e1b, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on latent human rights scoress"}
out_mfx_e1b <- mfx_e1d_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

@@

```{r plot-mfx-e2ab, fig.width=7, fig.height=4.5, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on the probability of specific levels of political terror and predicted latent human rights values"}
out_mfx_e2a <- mfx_e2a_pts %>% 
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(effect1__, effect2__, plot_var_nice) %>% 
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  mutate(fake_facet_title = "Political Terror Scale") %>% 
  ggplot(aes(x = effect1__, y = estimate__, color = effect2__, fill = effect2__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, color = NA) +
  geom_line(size = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.75, direction = -1) +
  scale_fill_viridis_d(option = "rocket", begin = 0.1, end = 0.9, direction = -1) +
  labs(x = "Civil society repression",
       y = "Predicted probability of category",
       color = "Political Terror Scale", fill = "Political Terror Scale",
       caption = "(x-axis reversed; greater repression on the right)") +
  facet_wrap(vars(fake_facet_title)) +
  theme_ngo() +
  theme(legend.box.margin = margin(t = 0.5, unit = "lines"))

out_mfx_e2b <- mfx_e2d_lhr %>% 
  mutate(fake_facet_title = "Latent human rights values") %>% 
  ggplot(aes(x = effect1__, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, 
              fill = "#FF4136", color = NA) +
  geom_line(size = 1, color = "#FF4136") +
  scale_x_reverse() +
  labs(x = "Civil society repression",
       y = "Predicted latent human rights value",
       caption = "(x-axis reversed; greater repression on right\nhigher values in y-axis show less violence and better protection of rights)") +
  facet_wrap(vars(fake_facet_title)) +
  theme_ngo()

@@

```{r correct-not-correct}
# These numbers come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html#Prediction_diagnostics
pts_pred_table <- tribble(
  ~Prediction, ~Baseline, ~total, ~advocacy, ~entry, ~funding, ~repression,
  "Wrong",     108,       108,    110,       108,    109,      111,
  "Correct",   377,       377,    375,       377,    376,      374
) %>% 
  rename("Total legal barriers included" = total,
         "Barriers to advocacy included" = advocacy,
         "Barriers to entry included" = entry,
         "Barriers to funding included" = funding,
         "Civil society repression included" = repression)

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html
tbl_rmse <- tribble(
  ~`Latent human rights values`,  ~RMSE, ~`% change from baseline`,
                  "Baseline", 0.1861944,                       "—",
      "Total legal barriers", 0.1872427,                   "0.56%",
      "Barriers to advocacy", 0.1871498,                   "0.51%",
         "Barriers to entry", 0.1862244,                   "0.02%",
       "Barriers to funding", 0.1879829,                   "0.96%",
  "Civil society repression", 0.1856543,                  "-0.29%"
) %>% 
  mutate(RMSE = round(RMSE, 4))

@@

When looking at the overall latent human rights values in a country, including de jure and de facto repression does very little to improve predictive accuracy. As Table \@ref(tab:tbl-rmse) shows, accounting for de facto civil society repression only reduces the RMSE by 0.29% compared to the baseline, while accounting for de jure laws worsen the models' predictive power slightly. While these gains in predictive power appear negligible, they are somewhat comparable with the findings of previous work. For instance, in their work on the effect of the murder of journalists on predicting future political terror, @GohdesCarey:2017 find that accounting for violence against the media adds five correctly predicted cases of PTS scores and improves predictions in seven cases.

>>>>

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
published: "Conditionally accepted at the *Journal of Human Rights*"
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while de jure anti-NGO laws provide little information in predicting future repression, their patterns of implementation—or de facto civil society repression—predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "8,343"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

@@

  # Load general data
  canary <- tar_read(panel)
  tar_load(vdem_clean)

  # Load Civicus stuff
  tar_load(c(civicus_map_data, civicus_clean))

  # Load models
  tar_load(c(m_lhr_total, m_lhr_advocacy, m_lhr_v2csreprss))

  # Load marginal effects
  tar_load(c(mfx_e1a_pts, mfx_e1b_lhr, mfx_e2a_pts, mfx_e2b_lhr))
})
```

@@

```{r civicus-map, fig.width=6.5, fig.height=3.1, fig.asp=NULL, fig.cap="2020 CIVICUS Monitor civic space ratings", fig.pos="h"}
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_viridis_d(option = "rocket", direction = -1,
                       na.translate = FALSE, name = "Civic space") +
  theme_void(base_size = 9, base_family = "Inter Light") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

@@

Simply counting these laws may therefore miss the effect of their de facto implementation. Some governments may even pass laws to create a chilling effect on these organizations. By providing a legal threat, they could compel these organizations to self-censor without even needing to implement these laws in any systematic or meaningful capacity. These laws, especially during moments or crises or contention, "highlight to the wider NGO community-and citizenry-that the government is willing and capable of using the law to repress dissent" [@Brechenmacher:2017, 94].

@@

```{r repression, fig.width=5.5, fig.height=4, fig.asp=NULL, fig.cap="Key measures of repression over time"}
pts_plot <- canary %>% 
  filter(!is.na(PTS_factor)) %>% 
  group_by(PTS_factor, year) %>% 
  summarize(count = n()) %>% 
  mutate(fake_facet_title = "Political Terror Scale") %>% 
  ggplot(aes(x = year, y = count, color = PTS_factor)) +
  geom_line(size = 1) +
  scale_color_viridis_d(option = "rocket", end = 0.85, direction = -1) +
  coord_cartesian(xlim = c(1990, 2015)) +
  guides(color = guide_legend(title = NULL)) +
  labs(y = "Number of countries", x = NULL, caption = "(Higher levels = more repression)") +
  theme_ngo() +
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines"),
        plot.caption = element_text(margin = margin(t = -1, unit = "lines"))) +
  facet_wrap(vars(fake_facet_title))

@@

```{r plot-mfx-e1a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on the probability of specific levels of political terror"}
mfx_e1a_pts %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  # Collapse 3-5 to just one level, since they're all tiny
  mutate(effect2__ = fct_collapse(effect2__, "Level 3+" = 
                                    c("Level 3", "Level 4", "Level 5"))) %>% 
  group_by(x_rounded, effect2__, plot_var_nice) %>% 
  slice(1) %>% 
  # Add up these values for the collapsed 3-5 results
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = effect2__)) +
  geom_line(aes(group = effect2__), size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  # position = position_dodge(width = 0.2),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.85, direction = -1) +
  guides(color = guide_legend(override.aes = list(size = 0.5, linetype = 0))) +
  labs(x = "Count of legal barriers\n",
       y = "Predicted probability of level",
       color = "Predicted Political Terror Scale", fill = "Predicted Political Terror Scale",
       title = "Political Terror Scale") + 
  facet_wrap(vars(plot_var_nice), scales = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())
```

@@

```{r plot-mfx-e1b, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on latent human rights scoress"}
out_mfx_e1b <- mfx_e1b_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

@@

```{r plot-mfx-e2ab, fig.width=7, fig.height=4.5, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on the probability of specific levels of political terror and predicted latent human rights values"}
out_mfx_e2a <- mfx_e2a_pts %>% 
  mutate(effect2__ = fct_collapse(effect2__, "Level 3+" = 
                                    c("Level 3", "Level 4", "Level 5"))) %>% 
  group_by(effect1__, effect2__, plot_var_nice) %>% 
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  mutate(fake_facet_title = "Political Terror Scale") %>% 
  ggplot(aes(x = effect1__, y = estimate__, color = effect2__, fill = effect2__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, color = NA) +
  geom_line(size = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.75, direction = -1) +
  scale_fill_viridis_d(option = "rocket", begin = 0.1, end = 0.9, direction = -1) +
  labs(x = "Civil society repression",
       y = "Predicted probability of level",
       color = NULL, fill = NULL,
       caption = "(x-axis reversed; greater repression on the right)") +
  facet_wrap(vars(fake_facet_title)) +
  theme_ngo() +
  theme(legend.box.margin = margin(t = 0.5, unit = "lines"))

out_mfx_e2b <- mfx_e2b_lhr %>% 
  mutate(fake_facet_title = "Latent human rights values") %>% 
  ggplot(aes(x = effect1__, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, 
              fill = "#FF4136", color = NA) +
  geom_line(size = 1, color = "#FF4136") +
  scale_x_reverse() +
  labs(x = "Civil society repression",
       y = "Predicted latent human rights value",
       caption = "(x-axis reversed; greater repression on right\nhigher values in y-axis show less violence and better protection of rights)") +
  facet_wrap(vars(fake_facet_title)) +
  theme_ngo()

@@

```{r correct-not-correct}
# These numbers come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html#Prediction_diagnostics
pts_pred_table <- tribble(
  ~Prediction, ~Baseline, ~total, ~advocacy, ~entry, ~funding, ~repression,
  "Wrong",     109,       107,    110,       108,    108,      111,
  "Correct",   376,       378,    375,       377,    376,      374
) %>% 
  rename("Total legal barriers included" = total,
         "Barriers to advocacy included" = advocacy,
         "Barriers to entry included" = entry,
         "Barriers to funding included" = funding,
         "Civil society repression included" = repression)

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html
baseline_rmse <- 0.1862156
tbl_rmse <- tribble(
  ~`Latent human rights values`, ~RMSE,
                  "Baseline",    baseline_rmse,
      "Total legal barriers",    0.1872181,
      "Barriers to advocacy",    0.1872011,
         "Barriers to entry",    0.1862635,
       "Barriers to funding",    0.1879768,
  "Civil society repression",    0.1857038
) %>% 
  mutate(`% change from baseline` = (RMSE - baseline_rmse) / baseline_rmse,
         `% change from baseline` = percent(`% change from baseline`, accuracy = 0.01)) %>% 
  mutate(`% change from baseline` = recode(`% change from baseline`, "0.00%" = "—")) %>% 
  mutate(RMSE = round(RMSE, 4))

@@

When looking at the overall latent human rights values in a country, including de jure and de facto repression does very little to improve predictive accuracy. As Table \@ref(tab:tbl-rmse) shows, accounting for de facto civil society repression only reduces the RMSE by 0.27% compared to the baseline, while accounting for de jure laws worsen the models' predictive power slightly. While these gains in predictive power appear negligible, they are somewhat comparable with the findings of previous work. For instance, in their work on the effect of the murder of journalists on predicting future political terror, @GohdesCarey:2017 find that accounting for violence against the media adds five correctly predicted cases of PTS scores and improves predictions in seven cases.
